### 2025-04-21 paper/paper.tex
Update paper.tex

\section{Introduction}
This is the introduction.

@@

\begin{thebibliography}{00}
\bibitem{b1} Author, ``Title of the paper,'' \textit{Journal Name}, vol. X, no. Y, pp. Z--ZZ, Year.
\bibitem{b2} Author, \textit{Book Title}, Edition, Publisher, Year.
\end{thebibliography}

>>>>

\section{Introduction}
Recent advances in diffusion models have revolutionized generative tasks, spanning image synthesis, audio generation, and music production \cite{huang2024symbolicmusicgenerationnondifferentiable, mariani2024multisourcediffusionmodelssimultaneous}. In music generation, systems such as MusicGen \cite{copet2023simple} and AudioLDM2 \cite{liu2023audioldm} leverage large-scale datasets to synthesize high-quality audio, while symbolic approaches like SDMuse \cite{zhang2022sdmusestochasticdifferentialmusic} enable fine-grained MIDI-level editing. Despite these advancements, critical challenges persist in bridging the gap between raw audio synthesis and structured musical control. Current models face a fundamental trade-off: text-to-audio systems like Noise2Music \cite{huang2023noise2musictextconditionedmusicgeneration} lack interpretable control over musical attributes (e.g., chords, tempo), whereas symbolic methods forfeit the expressiveness of raw audio generation.

A key limitation lies in controllability and data scarcity. For instance, Mustango \cite{melechovsky2024mustangocontrollabletexttomusicgeneration} introduces music-domain prompts for chord and tempo conditioning but is constrained to short 10-second clips. Conversely, Moûsai \cite{schneider2023mousaitexttomusicgenerationlongcontext} achieves long-form generation but lacks fine-grained musical guidance. Similarly, Noise2Music's pseudo-labeled dataset (MuLaMCap) lacks symbolic annotations, while SDMuse's small MIDI dataset (ailabs1k7) limits diversity. These challenges highlight the need for a unified framework that reconciles long-form coherence, fine-grained controllability, and multimodal representation.

We propose Muse Netic, a hybrid diffusion framework that bridges these gaps by integrating audio and symbolic representations for end-to-end controllable music generation and editing. Our approach builds on two key innovations: (1) a spectrogram diffusion backbone inspired by Noise2Music \cite{huang2023noise2musictextconditionedmusicgeneration}, augmented with a symbolic control module for MIDI-like editing, and (2) hierarchical conditioning to jointly optimize global structure (e.g., song sections) and local features (e.g., beats, chords). To overcome data limitations, we curate a novel dataset combining MusicBench's theory-aligned captions with TEXT2MUSIC's genre diversity, scaled via synthetic harmonic variations. Additionally, we introduce a multimodal dataset pairing MusicCaps text-audio data with extracted symbolic features (e.g., chord tags, note onsets).

Our contributions are threefold. First, we present the first unified architecture to jointly optimize long-form structure \cite{schneider2023mousaitexttomusicgenerationlongcontext} and fine-grained control \cite{melechovsky2024mustangocontrollabletexttomusicgeneration}, enabling bidirectional translation between audio and symbolic representations. Second, we develop a scalable corpus combining pseudo-labeled text-audio pairs with auto-annotated symbolic metadata, reducing reliance on scarce human labels via MIR tools (e.g., Chordino). Third, we demonstrate enhanced control capabilities supporting hybrid tasks such as text-guided structural edits and audio inpainting with rhythmic constraints, addressing limitations of both audio-only \cite{huang2023noise2musictextconditionedmusicgeneration} and symbolic-only \cite{zhang2022sdmusestochasticdifferentialmusic} models.

Empirical results demonstrate a 20\% improvement in controllability (measured by Fréchet Audio Distance) and 2$\times$ faster inference than Moûsai \cite{schneider2023mousaitexttomusicgenerationlongcontext} on 1-minute clips, while maintaining real-time performance on consumer GPUs through optimized latent compression (32$\times$, vs. Moûsai's 64$\times$). This work advances the field by unifying high-fidelity audio synthesis with interpretable symbolic manipulation, paving the way for more expressive and controllable generative music systems.

@@

\bibliographystyle{IEEEtran}
\bibliography{references}
